---
title: 'Assignment #5'
author: "Richard Viebrock and David Segan"
date: "11/27/2019"
output: html_document
---

```{r setup, include = FALSE}
knitr::opts_chunk$set(echo = FALSE, message = FALSE, warning = FALSE)
```

### **Introduction**

A short (7 - 8 sentence) introduction that helps the reader understand the purpose/motivation of the research, and previews the content of the report. Add an image of the Pacific giant salamander, making sure to give appropriate credit to the author / creator) in the image figure caption.

### **Data and Methods**

A brief ‘Data and Methods’ section (5 - 6 sentences) summarizing the data and how it was analyzed. Briefly describe the variables being studied in this report. Include the types of statistical tests performed, significance level used, and software (with version) for analysis. Include a map of the area. You can use an already published map of the area, making sure to give appropriate credit to the author / creator) in the map figure caption.

```{r}

# ------------------------
# Attach Relevant Packages
# ------------------------

library(tidyverse)
library(janitor)
library(directlabels)
library(DT)
library(kableExtra)
library(ggbeeswarm)
library(car)
library(wesanderson)

```

### **Results**

#### **Part A:** Salamander Counts by Section

```{r}

# --------------------------------
# Read in data and filter for DITE 
# --------------------------------

mack_creek <- read_csv("mack_creek_vertebrates.csv") %>% 
  clean_names() %>% 
  filter(species == "DITE")

```

<center>
```{r, fig.width=6, fig.height=6, fig.cap = "**Figure 1**: *Insert caption here.*"}

# --------------------------------
# Results A: Create DF for salamander count graph
# --------------------------------

mack_creek_count <- mack_creek %>% 
  filter(section %in% c("CC", "OG")) %>% 
  mutate(section_name = if_else(section == "CC", "Clear Cut", "Old Growth")) %>% 
  select(year, section_name) %>% 
  group_by(year, section_name) %>% 
  summarize(count = n())

# -------------------------------------------
# Use ggplot to create salamander count graph
# -------------------------------------------

ggplot(data = mack_creek_count, aes(x = year, y = count, group = section_name))+
  geom_line(aes(color = section_name), size = 1, show.legend = FALSE)+
  geom_point(color = "gray30")+
  geom_dl(aes(label = section_name, color = section_name), method = list(dl.combine("last.points"), cex = 0.85))+
  scale_color_manual(values = wes_palette("Cavalcanti1"))+
  scale_x_continuous(expand = c(0, 0),
                   limits = c(1993, 2021),
                   breaks = seq(1993, 2018, by = 5))+
  scale_y_continuous(expand = c(0, 0),
                     limits = c(0, 400),
                     breaks = seq(0, 400, by = 100))+
  labs(x = "Year",
       y = "Count",
       title = "Salamander Counts by Section: 1993 - 2017")+
  theme_minimal()

```
</center>

#### Analysis of Results:
- There is little visual difference between salamander counts in old growth and clear cut areas between 1993 and 2017. 
- Both old growth and clear cut areas experienced significant decline in salamander population between 2013 and 2015.

&nbsp;

#### **Results B:** 2017 Salamander Counts by Channel Classification

**Figure 2:** *Insert caption here*
```{r}

# ---------------------------------------------------------------------
# Results B: Create DF for table (channel classification & forest type)
# ---------------------------------------------------------------------

mack_c_table <- mack_creek %>% 
  filter(section %in% c("OG", "CC")) %>% 
  mutate(section_name = if_else(section == "CC", "Clear Cut", "Old Growth")) %>% 
  select(year, section_name, unittype) %>%
  filter(year == 2017,
         unittype %in% c("C", "P", "SC")) %>% 
  group_by(section_name, unittype) %>% 
  summarize(count = n()) %>% 
  pivot_wider(names_from = unittype, values_from = count) %>% 
  rename(Cascades = "C",
         Pool = "P",
         "Side Channel" = "SC",
         Section = "section_name")

mack_c_props <- mack_c_table %>% 
  adorn_percentages(denominator = "row") %>% 
  adorn_pct_formatting(digits = 1) %>% 
  adorn_ns(position = "front")

# ------------------------------------
# Create contingency table using kable
# ------------------------------------

kable(mack_c_props) %>% 
  kable_styling()

```

&nbsp;

#### **Results C:** Analysis of Salamander Counts by Channel Classification
```{r}

# ----------------------------------------------------
# Results C: Use chi-squared to determine independence
# ----------------------------------------------------

chi_mack <- mack_c_table %>%
  ungroup() %>% 
  select(-Section)

chi_mack_test <- chisq.test(chi_mack)

# Regarding the locations where salamanders were observed in Mack Creek in 2017, is there a significant effect of forest condition (old growth or clear cut) on where in the channel salamanders are found (channel classification)? Report your statistical findings in text, in the context of the actual counts / proportions you present in the table. 
```

- Chi Squared test of independence p-value: `r round(chi_mack_test$p.value, 2)`. 
- There is not a significant effect of forest condition on where channel salamanders are found. 

&nbsp;

#### **Results D:** Pacific Giant Salamander Weights by Section
<center>
```{r, fig.width=6, fig.height=6, fig.cap = "**Figure 1**: *Insert caption here.*"}

# ---------------------------------------------
# Results D: Create data frame to compare means
# ---------------------------------------------

mack_c_weights_graph <- mack_creek %>% 
  select("year", "section", "weight") %>% 
  filter(section %in% c("OG", "CC"),
         year == 2017)

mack_c_summary <- mack_c_weights_graph %>% 
  group_by(section) %>% 
  summarize(
    mean_weight = mean(weight, na.rm = TRUE),
    sd_weight = sd(weight, na.rm = TRUE))

# ---------------------------------------
# Use ggbeeswarm to visually compare means
# ---------------------------------------

ggplot()+
  geom_jitter(data = mack_c_weights_graph, 
                aes(x = section, y = weight, color = section),
                size = 1.5,
                alpha = 0.6,
                show.legend = FALSE) +
  scale_color_manual(values = wes_palette("Cavalcanti1"))+
  geom_point(data = mack_c_summary,
             aes(x = section, y = mean_weight),
             color = "black",
             size = 2.25) +
  geom_errorbar(data = mack_c_summary,
                aes(x = section,
                    ymin = mean_weight - sd_weight,
                    ymax = mean_weight + sd_weight),
                width = 0.1,
                size = 1.15,
                color = "black")+
  scale_x_discrete(labels = c("Clear Cut", "Old Growth"))+
  labs(x = "Section",
       y = "Weight",
       title = "2017 Means Comparison: Old Growth vs Clear Cut")+
  theme_minimal()

# Make sure that the beeswarm and the means are lined up appropriately. How do you make sure? Visual Check?
```
</center>
```{r}

# --------------------------------------------------
# Conduct a means comparison using a 2-sample t-test
# --------------------------------------------------

 mack_weights_cc <- mack_c_weights_graph %>% 
   filter(section == "CC") %>% 
   pull(weight)

 mack_weights_og <- mack_c_weights_graph %>% 
   filter(section == "OG") %>% 
   pull(weight)

 mack_means_comparison <- t.test(mack_weights_og, mack_weights_cc)

```

#### Analysis of Results: 
- Welch's 2-sample t-test p-value: `r round(mack_means_comparison$p.value, 3)`
- There is not a significant difference in weights between giant salamanders in clear cut and old growth sections.

&nbsp;

#### **Results E:** Means Comparison of Salamander Weight by Channel Classification
##### Visual Comparison of Salamander Weight by Channel Classification

```{r}

# ---------------------------------------------------
# Results E: Create data frame to compare by unittype
# ---------------------------------------------------

 mack_pcs_weights <- mack_creek %>% 
  select("year", "unittype", "weight", "section") %>% 
   filter(unittype %in% c("C", "P", "SC"),
          year == 2017,
          section %in% c("CC", "OG")) %>% 
  mutate(unittype = ifelse(unittype == "C", "Cascades",
                           ifelse(unittype == "P", "Pool", "Side Channel")))
 
 mack_pcs_summary <- mack_pcs_weights %>% 
   group_by(unittype) %>% 
   summarize(
     mean_weight = mean(weight, na.rm = TRUE),
     sd_weight = sd(weight, na.rm = TRUE),
     sample_size = n(),
     se_weight = sd(weight, na.rm = TRUE) / sqrt(n()),
     var_weight = var(weight, na.rm = TRUE))

```
<center>
```{r}

# -----------------------------------
# Make beeswarm plot to compare means
# -----------------------------------

ggplot()+
   geom_beeswarm(data = mack_pcs_weights,
                 aes(x = unittype, y = weight, color = unittype),
                 size = 1,
                 alpha = 0.6,
               show.legend = FALSE) +
   scale_color_manual(values = wes_palette("Darjeeling1"))+
   geom_point(data = mack_pcs_summary,
              aes(x = unittype, y = mean_weight),
              color = "black",
              size = 2)+
   geom_errorbar(data = mack_pcs_summary,
                 aes(x = unittype,
                     ymin = mean_weight - sd_weight,
                     ymax = mean_weight + sd_weight),
                 width = 0.1,
                 color = "black") +
   scale_x_discrete(labels = c("Cascade", "Pool", "Side Channel"))+
  labs(x = "Unit Type",
       y = "Weight (grams)",
       title = "Means Comparison by Channel Clasification")+
  theme_light()

 # Make sure that data is lined up correctly so that you don't have a mean in the wrong place. If the mack_pcs_weights is in the same order as mack_pcs_summary, everything works

```
</center>

&nbsp;

<center>
```{r}

# ------------------------------------
# Make a QQ plot to determine normalcy
# ------------------------------------

ggplot(data = mack_pcs_weights, aes(sample = weight))+
  geom_qq(aes(color = unittype), show.legend = FALSE)+
  scale_color_manual(values = wes_palette("Darjeeling1"))+
  facet_wrap(~unittype)+
  labs(x = "Theoretical",
       y = "Samples",
       title = "Analysis of Data Spread Normalcy")

# ggplot(data = mack_pcs_weights, aes(x = weight))+
#  geom_histogram(aes(fill = unittype), 
#                 color = "black", 
#                 show.legend = FALSE)+
#  scale_fill_manual(values = wes_palette("Darjeeling1"))+
#  facet_wrap(~unittype)+
#  labs(x = "Weight",
#       y = "Count")

# ------------------------------------------------
# Conduct a Levene's test to ensure equal variance
# ------------------------------------------------

pcs_levene_test <- leveneTest(weight ~ unittype, data = mack_pcs_weights, na.rm = TRUE)

# ----------------------------------------------------
# Conduct a chi-squared test to check for independence?
# ----------------------------------------------------

chi_pcs_mack <- mack_pcs_weights %>%
  group_by(unittype) %>% 
  summarize(weight = mean(weights))

```
</center>

#### Levene's Test for Equal Variance
- Levene's Test Result: With a p-value of 0.09, this is not significant and we retain the null hypothesis of equal variances. 

```{r}

# ----------------------------------------
# Conduct a one-way ANOVA to compare means
# ----------------------------------------

mack_aov <- aov(weight ~ unittype, data = mack_pcs_weights)

# ----------------------------------------
# Conduct Tukey's HSD for post-hoc testing
# ----------------------------------------

# TukeyHSD(mack_aov)

```

#### One-way ANOVA & Post-Hoc Testing:
- ANOVA Result: `r (mack_aov$p.value)`
- Tukey's HSD Result:

#### Analysis of Results:
- Interesting trend 1
- Interesting Trend 2
- Interesting Trend 3

### **Summary**
A brief summary of the major findings (pick 3 - 4) from your mini-report. A bullet-pointed list is fine, but the findings should be well-written, responsible (don’t overstate your findings), and refer to outcomes (e.g. figures, tables) in the Results section

### **References**